LME-MPM applied to quasi-brittle fracture.

### 2020-06-30 2020_EFM_MPM_Eigensoftening.tex

Figure \ref{fig:Reactions-cracked-panel-mode-I} clearly shows that the
\acrshort{lme}$_{\gamma = 3.5}$ solution has a reaction peak value of 0.03 $[M] [L]
[T]^{-2}$ for a imposed displacement 0.015 $[L]$ which agrees with the
analytical solution \cite{Pandolfi_2012}. It also shows the presence
of wiggles in the reaction-displacements curve, in both loading and
post-failure stages, when linear interpolation technique is used. In
contrast, \acrshort{lme} simulation does not exhibit these spurious
oscillations and remains linear until the failure. About the strength,
linear interpolation produces stiffer results than \acrshort{lme}. This can be attributed to imprecision in the
stress field due to grid-crossing phenomena. Alternative to
  \acrshort{lme}, \acrshort{ugimp} shape
functions can be employed to mitigate grid-crossing instabilities.
Figure \ref{fig:Reactions-cracked-panel-mode-I} support this since it
does not exhibits it when particles cross element borders as occur
with linear interpolation. Unfortunately the extended support of this
kind of shape function interfere with nodes in both sides of the notch,
producing an undesired ``tie-effect''. This involves in a violation of
the free border boundary condition in the notch. Possible solutions to this effect
are the suppression of the elements in the notch, within a CRAMP-like
method where two pairs of nodes are employed, or reducing the mesh size. With \acrshort{lme}
approximants these artefacts are not required since it is possible to control the support
of the shape function reducing the grid-crossing and avoiding the
``tie-effect'' in the notch at the same time.\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figure-Reactions-Mode-I}
  \caption{Evolution of the reaction forces plotted \textit{versus}
    the imposed displacement for linear interpolation technique and
    the \acrshort{lme} approximants.}
  \label{fig:Reactions-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Additionally, Figure \ref{fig:Stress-cracked-panel-mode-I}
remark clearly the difference between linear interpolation and
\acrshort{lme}, since numerical instability becomes rather significant
when particles crosses the boundary of the element in the linear
case, which is observed at 60 seconds. Although, apparently, the
differences in the obtained peaks are not extremely significant, it
can be owing to the aforementioned hookean material
employed~\cite{Zhang_EE_2020}. When more sophisticated constitutive
models will be employed, severe inaccuracies may appear in the
stress field that could affect dramatically to the final 
result. Focusing on the post failure behavior, it can be seen how
\acrshort{lme} produces soft stress field evolution even after
breaking. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[t = 25]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0100-mode-I.png}}
\subfigure[t = 60]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0250-mode-I.png}}
\subfigure[t = 100]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0398-mode-I.png}}
\caption{Evolution of the stress tensor magnitude for linear
  interpolation (pictures in the left side), and \acrshort{lme}
  (pictures in the right side). Both simulations performed with an
  eigenerosion algorithm.}
\label{fig:Stress-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
An important consideration regarding the presence of oscillations once
both parts of the panel are separated is observed in Figure
\ref{fig:Reactions-cracked-panel-mode-I}. These phenomena should not be
attributed to the eigenerosion algorithm since the fracture process is
over. They are due to the dynamic nature of the solver: once the
energy is released dinamically by the crack, the waves propagate
along the domain.

>>>>

Figure \ref{fig:Reactions-cracked-panel-mode-I} clearly shows that the
\acrshort{lme}$_{\gamma = 3.5}$ solution has a reaction peak value of 0.03 $[M] [L]
[T]^{-2}$ for a imposed displacement 0.015 $[L]$ which agrees with the
analytical solution \cite{Pandolfi_2012}. It also shows the presence
of wiggles in the reaction-displacements curve, in both loading and
post-failure stages, when linear interpolation technique is used. In
contrast, \acrshort{lme} simulation does not exhibit these spurious
oscillations and remains linear until the failure. About the strength,
linear interpolation produces stiffer results than \acrshort{lme}. This can be attributed to imprecision in the
stress field due to grid-crossing phenomena. Alternative to
  \acrshort{lme}, \acrshort{ugimp} shape
functions can be employed to mitigate grid-crossing instabilities.
Figure \ref{fig:Reactions-cracked-panel-mode-I} supports this fact since it
does not exhibit instabilities when particles cross element borders meanwhile these instabilities are found with linear interpolation. Unfortunately, the extended support of this
kind of shape function interfere with nodes in both sides of the notch,
producing an undesired ``tie-effect''. This involves in a violation of
the free border boundary condition in the notch. Possible solutions to this effect
are the suppression of the elements in the notch, within a CRAMP-like
method where two pairs of nodes are employed, or reducing the mesh size. With \acrshort{lme}
approximants these artefacts are not required since it is possible to control the support
of the shape function reducing the grid-crossing and avoiding the
``tie-effect'' in the notch at the same time.\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \includegraphics[width=0.8\textwidth]{Figure-Reactions-Mode-I}
  \caption{Evolution of the reaction forces plotted \textit{versus}
    the imposed displacement for linear interpolation technique and
    the \acrshort{lme} approximants.}
  \label{fig:Reactions-cracked-panel-mode-I}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Additionally, Figure \ref{fig:Stress-cracked-panel-mode-I}
remark clearly the difference between linear interpolation and
\acrshort{lme}, since numerical instability becomes rather significant
when particles crosses the boundary of the element in the linear
case, which is observed at 60 seconds. Although, apparently, the
differences in the obtained peaks are not extremely significant, it
can be owing to the aforementioned hookean material
employed~\cite{Zhang_EE_2020}. When more sophisticated constitutive
models will be employed, severe inaccuracies may appear in the
stress field that could affect dramatically to the final 
result. Focusing on the post failure behavior, it can be seen how
\acrshort{lme} produces soft stress field evolution even after
breaking. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subfigure[t = 25]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0100-mode-I.png}}
\subfigure[t = 60]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0250-mode-I.png}}
\subfigure[t = 100]{\includegraphics[width=0.73\textwidth]{./Figure-Stress-0398-mode-I.png}}
\caption{Evolution of the stress tensor magnitude for linear
  interpolation (pictures in the left side), and \acrshort{lme}
  (pictures in the right side). Both simulations performed with an
  eigenerosion algorithm.}
\label{fig:Stress-cracked-panel-mode-I}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
An important consideration regarding the presence of oscillations once
both parts of the panel are separated is observed in Figure
\ref{fig:Reactions-cracked-panel-mode-I}. These phenomena should not be
attributed to the eigenerosion algorithm since the fracture process is
over. They are due to the dynamic nature of the solver: once the
energy is released dinamically by the crack, the waves propagate
along the domain.
